---
title: "2016 Presidential Election Debrief"
output: pdf_document
---

Package

```{r,inclue=FALSE}
library(XML)
library(xml2)
library(tidyr)
library(data.table)
library(readxl)
library(readxl)
```

#### 2016 Presidential Election Debrief Project

### STEP 1. DATA WRANGLING

## 1. 2016 Presidential Election results reported at the country level.
reference: 
1)http://www.stat.berkeley.edu/users/nolan/data/voteProject/2016_US_County_Le
vel_Presidential_Results.csv
2)https://github.com/tonmcg/County_Level_Election_Results_1216/blob/master/2016_US_County_Level_Presidential_Results.csv

```{r, include=FALSE}

election.2016 = read.csv(file = "http://www.stat.berkeley.edu/users/nolan/data/voteProject/2016_US_County_Level_Presidential_Results.csv", header = TRUE) # read 2016 presidential election file

```

#REBUILDING ELECTION 2016
  In Election 2016, the data of Alaska is inconsistent and 'X' column is needless, thus dismiss the unnecessary rows and column.
  Since in Election 2004 and 2008 do not have any information about other parties except Democratic and Republican, the votes need to be switched to total_votes, which is total of all parties with other parites, to total of only two parties, Democratic and Republican. Then, weit also needs to be changed to percentage, differentiation, and differentiation of percentage of votes for Democratic and Republican.
  Also, there are some counties which have 'Parish' and 'city' instead of 'County' in their names, so the modification is requred to 'County'.

```{r, include=FALSE}
election.2016 = election.2016[election.2016$state_abbr != "AK",] #The data of Alaska is inconsistent, so I dismiss it.

election.2016 = election.2016[, names(election.2016) != 'X'] # 'X' column is needless, so I dismiss it

election.2016$total_votes = election.2016$votes_dem+election.2016$votes_gop #In election 2004 and 2008 do not have any information about other parties, we decide to change total_votes, which is total of all parties with other parties, to total of only Democratic and Republican

election.2016$per_dem = election.2016$votes_dem/election.2016$total_votes #percentage of Democratic

election.2016$per_gop = election.2016$votes_gop/election.2016$total_votes #percentage of Republican

election.2016$diff = abs(election.2016$votes_dem - election.2016$votes_gop) #differentiation of Democratic and Republican

election.2016$per_point_diff = abs(election.2016$per_dem -
                                     election.2016$per_gop) 
#differentiation of percentage of Democratic and Republican

election.2016 = as.data.frame(sapply(election.2016,gsub,pattern="Parish",replacement="County")) #there are some counties which have Parish instead of County in their names, so change them to County

election.2016 = as.data.frame(sapply(election.2016,gsub,pattern="city",replacement="County")) #there are some counties which have city instead of County in their names, so change them to County

election.2016 = data.frame(lapply(election.2016, as.character), stringsAsFactors=FALSE) #change factors to character strings

```

##2. 2012 Presidential Election results reported at the country level.
reference: 
1)http://www.politico.com/2012-election/map/#/President/2012/ 2)http://www.stat.berkeley.edu/users/nolan/data/voteProject/countyVotes2012/xxx.xml
3)http://www.stat.berkeley.edu/users/nolan/data/voteProject/countyVotes2012/stateNames.txt 

#READING ELECTION 2012 FILE AND REBUILDING STATES NAME FILE 

```{r, include=FALSE}
states = read.table("http://www.stat.berkeley.edu/users/nolan/data/voteProject/countyVotes2012/stateNames.txt", sep = "\t") #read states name file

states = sub(" +", "", states$V1) #remove white space

states = states[states != "states" & states != "alaska"] #remove Alaska since it is inconsistent data

states_info = paste("http://www.stat.berkeley.edu/users/nolan/data/voteProject/countyVotes2012/",states,".xml", sep = "") #make list of URLs to store all URLs of states in xmlparse

states_doc = lapply(states_info, read_xml) #parse XML file of states

```

#BUILDING ELECTION 2012

#GETTING NUMBER OF VOTES FOR EACH PARTY

```{r,include=FALSE}
votes_dem1 = list()
votes_dem = list()

for(i in 1:50){
  votes_dem1[[i]] = xml_find_all(xml_root(states_doc[[i]]), '//tr[@class= "party-democrat race-winner" or @class = "party-democrat"]/td[@class = "results-popular"]')
  votes_dem[[i]] = abs(tidyr::extract_numeric(as.character(votes_dem1[[i]])))
} #make a list of number of votes for Democratic for each state

votes_dem = unlist(votes_dem)

votes_gop1 = list()
votes_gop = list()

for(i in 1:50){
  votes_gop1[[i]] = xml_find_all(xml_root(states_doc[[i]]), '//tr[@class= "party-republican race-winner" or @class = "party-republican"]/td[@class = "results-popular"]')
  votes_gop[[i]] = abs(tidyr::extract_numeric(as.character(votes_gop1[[i]])))
}

votes_gop = unlist(votes_gop)
votes_data = data.frame(votes_dem, votes_gop, 
                  stringsAsFactors = FALSE)       #make small dataframe for votes_dem and votes_gop.

total_votes = votes_data$votes_dem+votes_data$votes_gop #get total votes of Democratic and Republican parties

votes_data = data.frame(votes_data,total_votes, stringsAsFactors = FALSE) #merge votes_dem, votes_gop, and total_votes


```

#PERCENTAGE OF DEMOCRATIC AND REPUBLICAN // DIFFERENTIATION 

```{r, include=FALSE}
per_dem = votes_data$votes_dem/votes_data$total_votes #percentage of Democratic
per_gop = votes_data$votes_gop/votes_data$total_votes #percentage of Republican
diff = abs(votes_data$votes_dem - votes_data$votes_gop) #differentiation of Democratic and Republican
per_point_diff = abs(per_dem - per_gop) #differentiation of percentage of Democratic and Republican
```

#ABBREVIATION OF STATES FOR ELECTION 2012

```{r, include=FALSE}
abbreviation_state = c("AL", "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "FL", "GA", "HI","ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN","MS", "MO",  "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH","OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VT", "VA", "WA","WV", "WI", "WY") #make abbreviation of states from election 2016 to apply them to election 2012
state_abbr1 = list()
for(i in 1:50){
  state_abbr1[[i]] = rep(abbreviation_state[i],length(votes_dem1[[i]]))
}
state_abbr = unlist(state_abbr1) #make it unlist to merge to dataframe

```

#COUNTY NAME

```{r, include=FALSE}
county_name1 = list()
county_name = list()
for(i in 1:50){
  county_name1[[i]] = xml_find_all(xml_root(states_doc[[i]]), '//th[@class = "results-county"]')
  county_name1[[i]] = county_name1[[i]][-1] #remove first row whose name is "County"
  county_name[[i]] = as.character(stringr::str_sub(county_name1[[i]], 40, -64))
}

county_name = unlist(county_name)

county_name[county_name != "District of Columbia"] = paste0(county_name[county_name != "District of Columbia"],' County') #insert ' County' to end of all strings to make county_name same as the election 2016
```

#COUNTY ID

```{r, include=FALSE}
combined_fips1 = list()
combined_fips = list()
for(i in 1:50){
  combined_fips1[[i]] = xml_find_all(xml_root(states_doc[[i]]), '//tbody')
  combined_fips[[i]] = substring(combined_fips1[[i]], regexpr('county', combined_fips1[[i]]) + 6, regexpr('">\n', combined_fips1[[i]])-1)
  combined_fips[[i]] = str_extract(combined_fips[[i]], "[:digit:]+")
}  #make a list of attributes of id's of tbody to get combined_fips

combined_fips = unlist(combined_fips)
#remove needless word and make them unlist and numeric to merge into dataframe.
```

#MERGING by Combined_fips

```{r, include=FALSE}

election.2012 = cbind(votes_data, per_dem, per_gop, diff, per_point_diff, state_abbr, county_name, combined_fips) #merge all data of 2012 election
election.2012 = as.data.frame(sapply(election.2012,gsub,pattern="Saint",replacement="St.")) #there are some county which have Saint instead of St. in their names, so change them to St.
election.2012 = as.data.frame(sapply(election.2012,gsub,pattern="County County",replacement="County")) #there are some county which have "County County" instead of "County" in their names, so change them to "County"
election.2012$combined_fips = as.character(election.2012$combined_fips)


drops = c("state_abbr","county_name")
election.2012 = election.2012[, !(names(election.2012) %in% drops)]

election_2012_2016 = merge(election.2016, election.2012, by = c("combined_fips")) #merge election 2016 and 2012

setnames(election_2012_2016, old = c("votes_dem.x", "votes_gop.x", "total_votes.x", "per_dem.x", "per_gop.x", "diff.x", "per_point_diff.x", "votes_dem.y", "votes_gop.y", "total_votes.y", "per_dem.y", "per_gop.y", "diff.y", "per_point_diff.y"), 
         new = c("Clinton_2016", "Trump_2016", "Total_2016", "DemProp_2016", "RepProp_2016", "DiffVotes_2016", "DiffPer_2016", "Obama_2012", "Romney_2012", "Total_2012", "DemProp_2012", "RepProp_2012", "DiffVotes_2012", "DiffPer_2012"))

```

##3. 2008 Presidential Election results reported at the country level.
Author: Roozbeh Faghihi Moghdam / 25765752
reference:
1)http://www.stat.berkeley.edu/users/nolan/data/voteProject/countyVotes2008.xlsx

```{r, include=FALSE}

result_2008 = tempfile(fileext = ".xlsx")
download.file(url = "http://www.stat.berkeley.edu/users/nolan/data/voteProject/countyVotes2008.xlsx", destfile = result_2008, mode="wb")

#Reading Data
result_2008 <- sapply(excel_sheets(result_2008), read_excel, path = result_2008)

```
